Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-12-15 rebuttal.tex

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

Dear Editor, Dear Reviewers,

\vspace{2em}

We would like to thank both the reviewers and the editor for going through our
manuscript. As a consequence of your comments, the manuscript underwent
several significant changes, that we summarize hereafter.

While the reviews are generally positive (quoting one of the reviewer, ``this
is an excellent model for this kind of research, and the results are
promising''), we certainly take the two main concerns raised from the reviews:

\begin{itemize}
    \item The reporting of the studies is lacking
    \item Possible ethical concerns are not emphasized
\end{itemize}

This revision of our manuscript attempts to address both these concerns.

\section*{Reporting of the studies}

We have clarified our wording to only draw conclusion from factual measured
data or feedback from the parents/teachers/therapists (this is in particular
the case for the case-study 1),

We have added new quantitative data in the reporting of the case study 2, that
show a quantifiable improvement of handwriting. We want however to remain
cautious in our claims: handwriting acquisition is a complex cognitive skill.
While the child demonstrated clear improvements over the course of our
experiment, measuring the lasting effects of the robot-based remediation is
difficult (as many other individual and social factors may have impacted the
child's performance as well).

Finally, we have added the results of one more experiment (conducted during the
summer 2015), that involved 8 children with handwriting deficits. This
experiment focused on measuring how \emph{serious} the children were at
teaching the robot,

Other specific points were raised:

\begin{itemize}
    \item inclusions criteria should now be clearer,
    \item

\end{itemize}

\section*{Ethical concerns}

An explicit paragraph has been added to the discussion that states the possible
ethical concerns raised by this work. While the psychological impact of the
mentor-protégé relationship has been studied in the literature, it has been
done \emph{from the perspective of the protégé}. In our situation, the protégé
is the robot, and the child is the mentor: we could not find existing
literature discussing this situation, but would be nevertheless glad to further
expand the discussion if necessary.

\section*{Other comments}

One of the reviewers asked for clarification on a few other points:

\begin{itemize}

\item \emph{Did we include children who had difficulty with the dynamics of
turn-taking? If so, how did the system fare with those children?} In the case
the children did not pro-actively engaged into turn-taking, the facilitator (be
it one of the experimenter or the therapist) would suggest it.

\item \emph{The handwriting/letter generation is sufficiently described, but
how are the corresponding robot gestures generated?} We relied on NaoQI inverse
kinematics solver. This has been added to the text.

\item \emph{Was an interaction in which the child moves the hand of the NAO to
write the letter considered? Why did the authors choose the current interaction
model in which the child demonstrates on a tablet?} The Nao robot does not
feature force/torque sensing, which prevents most of the interesting direct
physical interaction like moving the hand of the robot. With appropriate
hardware, this interaction modality would be interesting to explore, though.

\item \emph{Could the measure of engagement plotted in Figure 7 be confounded
with the physically-active aspect of the hand-writing activity versus the
passive storytelling activity?} We have decided to replace this figure with an
objective metric (with-me-ness) that relies on the focus of attention of the
children to assess to which extend the child is into the task.

\item \emph{The paper would benefit from further explanation of the adult's
active role throughout the robot-child interaction} The manuscript hopefully
clarifies the role of the facilitator in the different studies.

\end{itemize}

\end{document}

### 2015-12-15 main.tex
Improved reporting on Thomas progress with simple satistical analysis

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}
    \caption{\small Demonstrations provided by Thomas for the number ``6'' (top row) and
        corresponding shapes generated by the robot. After eight demonstrations,
        Thomas decided that the robot's ``6'' was good enough, and went to
    another character: in that respect, he was the one leading the learning
process of the robot.}
    \label{learning_6_demos}
\end{figure}

@@

Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
shapes, the learning algorithm of the robot initially tended to converge towards
meaningless scribblings. We addressed this issue by having the robot to
\emph{refuse} allographs that were too far from the reference shape (the robt
would instead say \emph{``I'm
not sure to understand what you are drawing...''}), so that the child had to
pay good attention to what he would demonstrate to the robot. Also, to make
the robot's progresses evident, we modified the initialization step of the
learning algorithm to start with a roughly vertical stroke instead of a
deformed number (round 0 on Figure~\ref{learning_6_demos}).

@@

Since the robot's handwriting started from a simple primitive (a stroke), each
time Thomas succeeded to have his demonstration accepted by it, the robot's
improvement was clearly visible (as shown in Figure~\ref{Thomas_distances}).
This led to a self-rewarding situation that effectively supported Thomas'
engagement.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small Two metrics to assess the handwriting progresses: Euclidean
    distance in the subspace of the number dataset (top figure) or in
Cartesian space (bottom figure). Green lines represent the robot performance,
blue lines performance of the child. The round IDs correspond to the demonstrations
pictured on Figure~\ref{Thomas_demos}.}
    \label{Thomas_distances}
\end{figure}

\begin{figure}
    \centering
    \subfigure[Number ``2'']{
        \includegraphics[width=0.45\linewidth]{henry2}
    }
	\subfigure[Number ``5'']{
        \includegraphics[width=0.45\linewidth]{henry5}
    }
    \caption{\small Improvement of Thomas demonstrations for the numbers ``2''
        and ``5''. Thomas progressively took care of the demonstrations he was providing to the
    robot for those numbers. We used for this figure the same metric than the one
    used for the acceptance algorithm. Distances are normalized with respect to the
    biggest value. The dashed line correspond to the threshold of robot's
    acceptance.}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}

    \vspace{1em}

    \includegraphics[width=0.7\linewidth]{learning_6_distances}
    \caption{\small Demonstrations provided by Thomas for the number ``6'' (top row) and
        corresponding shapes generated by the robot. The plot beneath shows the distance
        (in the eigenspace of the shape) between Thomas demonstrations, the
        robot's attempts and our reference shape. After eight demonstrations,
        Thomas decided that the robot's ``6'' was good enough, and switched to
    another character. In that respect, he was the one leading the learning
process of the robot.}
    \label{learning_6_demos}
\end{figure}

@@

Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
shapes, the learning algorithm of the robot initially tended to converge towards
meaningless scribblings. We addressed this issue by having the robot to
\emph{refuse} allographs that were too far from the reference shape (the robt
would instead say \emph{``I'm
not sure to understand what you are drawing...''}), so that the child had to
pay good attention to what he would demonstrate to the robot. Also, to make
the robot's progresses evident, we modified the initialization step of the
learning algorithm to start with a roughly vertical stroke instead of a
deformed number (see the initial state on Figure~\ref{learning_6_demos}).

@@

Since the robot's handwriting started from a simple primitive (a stroke), each
time Thomas succeeded to have his demonstration accepted by it, the robot's
improvement was clearly visible (as shown in Figure~\ref{learning_6_demos}).
This led to a self-rewarding situation that effectively supported Thomas'
engagement.

\begin{figure}
    \centering
    \subfigure[Number ``2'']{
        \includegraphics[width=0.45\linewidth]{henry2}
    }
	\subfigure[Number ``5'']{
        \includegraphics[width=0.45\linewidth]{henry5}
    }
    \caption{\small Normalized distance between Thomas' demonstrations and
        reference allographs for the numbers ``2'' and ``5''. The horizontal
        dashed line correspond to the threshold for the robot to accept a
        demonstration. Thomas' progress is visible on these figures: we find a
        significant negative regression equation ($r=-0.023, F(1,19)=8.69,
        p<.02$, adjusted $R^2=.461$) for the number ``2'' (dotted red line),
        indicating that Thomas' shapes are getting closer to the reference. The
        regression is not significant for the number ``5'', but we can observe
        that after about 10 repetitions, all the demonstrations are deemed of
        acceptable quality by the robot.  }
